Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.

### 2024-06-14 manuscript/03_results.qmd
create supplemental smooth plots

| **Model**                          | Term                   | edf    | F     | p       |
|--------------------|---------------|-------------|-------------|-------------|
| **All Rhizaria Epipelagic**        | Salinity               | 2.0818 | 4.654 | \<0.001 |
| $R^2_{adj}$= 0.42                  | DO                     | 0.8272 | 0.945 | 0.0136  |
|                                    | Silica                 | 3.1206 | 15.75 | \<0.001 |
|                                    | NO3                    | 2.9970 | 6.268 | \<0.001 |
|                                    | Primary Productivity   | 1.8367 | 2.099 | \<0.001 |
|                                    | Particle Concentration | 0.9282 | 2.566 | \<0.001 |
| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6847 | 0.431 | 0.073   |
| $R^2_{adj}$= 0.406                 | Avg Mass Flux          | 1.5426 | 0.923 | 0.045   |
|                                    | Particle Concentration | 3.2957 | 20.89 | \<0.001 |
| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 1.6632 | 1.824 | 0.002   |
| $R^2_{adj}$= 0.603                 | Particle Concentration | 0.7694 | 0.662 | 0.027   |

: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column.

@@

| Model                                | Term                   | edf   | F     | p       |
|---------------|---------------|---------------|---------------|---------------|
| **Acantharea Epipelagic**            | Salinity               | 2.264 | 4.113 | \<0.001 |
| R2adj=0.53                           | O2                     | 2.579 | 6.712 | \<0.001 |
|                                      | Avg Mass Flux          | 4.810 | 22.81 | \<0.001 |
|                                      | Bacteria #/L           | 1.599 | 1.317 | 0.0117  |
|                                      | Particle Concentration | 0.853 | 1.158 | 0.0087  |
| **Acantharea Upper Mesopelagic**     | Avg C Flux             | 1.296 | 0.621 | 0.0439  |
| R2adj=0.231                          | Avg N Flux             | 1.619 | 1.283 | 0.0026  |
|                                      | Particle Concentration | 0.952 | 3.886 | \<0.001 |
| **Acantharea Lower Mesopelagic**     | Temperature            | 2.076 | 4.155 | \<0.001 |
| R2adj=0.509                          | Avg N Flux             | 1.494 | 1.216 | 0.0113  |
|                                      | Primary Productivity   | 0.766 | 0.648 | 0.0253  |
|                                      | Particle Concentration | 2.037 | 10.03 | \<0.001 |
| **Aulacanthidae Epipelagic**         | Salinity               | 0.792 | 0.757 | 0.0241  |
| R2adj=0.251                          | Avg N Flux             | 0.869 | 1.324 | 0.0018  |
|                                      | Primary Productivity   | 2.312 | 7.704 | \<0.001 |
|                                      | Particle Concentration | 2.008 | 2.200 | 0.0017  |
| **Aulacanthidae Upper Mesopelagic**  | Particle Concentration | 2.832 | 9.802 | \<0.001 |
| R2adj = 0.158                        |                        |       |       |         |
| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.622 | 1.991 | 0.002   |
| R2adj=0.298                          | Particle Concentration | 2.123 | 6.706 | \<0.001 |
| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.653 | 13.06 | \<0.001 |
| R2adj=0.2                            |                        |       |       |         |
| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 0.972 | 6.248 | \<0.001 |
| R2adj=.147                           |                        |       |       |         |
| **Castanellidae Epipelagic**         | Avg C Flux             | 1.462 | 0.816 | 0.0421  |
| R2adj = 0.124                        | Primary Productivity   | 0.771 | 0.665 | 0.0247  |
|                                      | Particle Concentration | 3.956 | 4.623 | \<0.001 |
| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.822 | 0.919 | 0.0183  |
| R2adj=.113                           | Particle Concentration | 0.970 | 6.208 | \<0.001 |
| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 1.773 | 4.873 | \<0.001 |
| R2adj=.133                           |                        |       |       |         |
| **Collodaria Epipelagic**            | Salinity               | 0.925 | 2.267 | \<0.001 |
| R2adj=0.16                           | O2                     | 2.015 | 2.217 | 0.002   |
|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.002   |
| **Foraminifera Epipelagic**          | Temperature            | 4.414 | 8.789 | \<0.001 |
| R2adj=0.445                          | O2                     | 1.603 | 1.287 | 0.0111  |
|                                      | Avg N Flux             | 2.512 | 3.396 | \<0.001 |
|                                      | Primary Productivity   | 0.824 | 0.920 | 0.0153  |
|                                      | Particle Concentration | 0.919 | 2.257 | \<0.001 |

: Taxa-specific generalized additive models for different regions of the water column.

>>>>

| **Model**                          | Term                   | edf    | F     | p       |
|--------------------|---------------|-------------|-------------|-------------|
| **All Rhizaria Epipelagic**        | Temperature            | 0.7950 | 0.492 | 0.0440  |
| $R^2_{adj}$= 0.48                  | Salinity               | 2.4117 | 5.113 | \<0.001 |
|                                    | Silica                 | 2.1511 | 8.601 | \<0.001 |
|                                    | Bacteria Concentration | 0.8870 | 1.544 | \<0.001 |
|                                    | Avg Mass Flux          | 1.4107 | 0.802 | 0.0211  |
|                                    | Primary Productivity   | 1.9280 | 2.736 | \<0.001 |
|                                    | Particle Concentration | 0.9151 | 2.136 | \<0.001 |
| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6702 | 0.404 | 0.0814  |
| $R^2_{adj}$= 0.40                  | Avg N Flux             | 2.9524 | 3.172 | \<0.001 |
|                                    | Particle Concentration | 3.5568 | 11.82 | \<0.001 |
| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 2.036  | 2.953 | \<0.001 |
| $R^2_{adj}$= 0.61                  | Particle Concentration | 3.819  | 27.43 | \<0.001 |

: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column. edf = estimated degrees of freedom, F = F-statistic, p = p-value.
Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.

@@

| Model                                | Term                   | edf   | F     | p       |
|--------------------------------------|------------------------|-------|-------|---------|
| **Acantharea Epipelagic**            | Temperature            | 0.868 | 0.380 | 0.0935  |
| $R^2_{adj}$=0.539                    | Salinity               | 2.349 | 4.646 |\<0.001  |
|                                      | O2                     | 2.501 | 5.493 |\<0.001  |
|                                      | Avg Mass Flux          | 4.815 | 22.59 |\<0.001  |
|                                      | Bacteria               | 1.608 | 1.360 | 0.00945 |
|                                      | Particle Concentration | 0.877 | 1.414 | 0.00449 |
| **Acantharea Upper Mesopelagic**     | Avg Mass Flux          | 1.866 | 1.220 | 0.0166  |
| $R^2_{adj}$=0.197                    | Avg N Flux             | 2.208 | 3.100 |\<0.001  |
|                                      | Primary Productivity   | 2.382 | 2.497 |\<0.001  |
|                                      | Particle Concentration | 0.724 | 0.521 | 0.0437  |
| **Acantharea Lower Mesopelagic**     | Temperature            | 1.958 | 2.986 |\<0.001  |
| $R^2_{adj}$=0.496                    | Avg F Flux             | 1.513 | 1.565 | 0.0037  |
|                                      | Primary Productivity   | 0.856 | 1.153 | 0.0045  |
|                                      | Particle Concentration | 1.661 | 8.667 |\<0.001  |
| **Aulacanthidae Epipelagic**         | Temperature            | 2.591 | 5.003 |\<0.001  |
| $R^2_{adj}$=0.394                    | Salinity               | 1.523 | 1.789 | 0.002   |
|                                      | RFU                    | 0.872 | 1.315 | 0.0042  |
|                                      | Bacteria               | 0.808 | 0.833 | 0.0160  |
|                                      | Primary Productivity   | 2.622 | 9.160 |\<0.001  |
|                                      | Particle Concentration | 1.937 | 2.751 |\<0.001  |
| **Aulacanthidae Upper Mesopelagic**  | Avg C Flux             | 1.675 | 1.408 | 0.0109  |
| $R^2_{adj}$=0.132                    | Particle Concentration | 2.308 | 2.476 | 0.0013  |
| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.476 | 1.418 | 0.0076  |
| $R^2_{adj}$=0.308                    | Particle Concentration | 2.457 | 7.756 | <0.001  |
| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.778 | 12.76 | <0.001  |
| $R^2_{adj}$=0.196                    |                        |       |       |         |
| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 2.012 | 6.306 |\<0.001  |
| $R^2_{adj}$=0.148                    |                         |      |       |         |
| **Castanellidae Epipelagic**         | Salinity               | 0.915 | 0.400 | 0.0935  |
| $R^2_{adj}$=0.228                    | RFU                    | 0.912 | 2.052 |\<0.001  |
|                                      | Si                     | 2.018 | 3.656 |\<0.001  |
|                                      | Bacteria               | 1.128 | 0.883 | 0.0170  |
|                                      | Particle Concentration | 2.278 | 4.028 |\<0.001  |
| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.734 | 0.551 | 0.0527  |
| $R^2_{adj}$=0.09                     | Particle Concentration | 0.965 | 5.102 |\<0.001  |
| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 3.051 | 6.768 |\<0.001  |
| $R^2_{adj}$=0.157                    |                        |       |       |         |
| **Collodaria Epipelagic**.           | Salinity               | 0.926 | 2.267 |\<0.001  |
| $R^2_{adj}$=0.16                     | O2                     | 2.015 | 2.217 |\<0.001  |
|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.0024  |
| **Foraminifera Epipelagic**          | Temperature            | 4.420 | 7.624 |\<0.001  |
| $R^2_{adj}$=0.543                    | O2                     | 0.799 | 0.788 | 0.023   |
|                                      | Bacteria.              | 0.000 | 0.000 | 0.523   |
|                                      | Avg N Flux             | 1.847 | 3.022 |\<0.001  |
|                                      | Primary Productivity   | 0.472 | 0.139 | 0.183   |
|                                      | Particle Concentration | 1.685 | 2.955 |\<0.001  |

: Taxa-specific generalized additive models for different regions of the water column.edf = estimated degrees of freedom, F = F-statistic, p = p-value.
Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.
